---
title: "Module 4: Enhancing Visualizations in R"
author: "Saswat Seth"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---
 

## Introduction

In this video, you'll learn how to change the aesthetics of your visuals to present your data in a more compelling way. Aesthetics help highlight key points in your data and communicate more effectively with stakeholders. 

An aesthetic is a visual property of an object in your plot. For instance, in a scatter plot, aesthetics include the size, shape, or color of your data points. You can display a point in different ways by adjusting its aesthetics.

## Basic Aesthetics

Let's review the code for plotting the relationship between body mass and flipper length using the penguins dataset. We start by mapping variables to the x and y axes using the `aes` function.

```{r}
library(ggplot2)

install.packages("palmerpenguins")

library(palmerpenguins)

data(penguins)
```

```{r}
# Basic scatter plot
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

### Adding Color

To differentiate between penguin species, we can map the `species` variable to the `color` aesthetic:

```{r}
# Aesthetic - color
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point()
```

### Adding Shape

Similarly, you can use the `shape` aesthetic to represent different penguin species with distinct shapes:

```{r}
# Aesthetic - shape
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, shape = species)) +
  geom_point()
```

### Combining Color and Shape

Combining multiple aesthetics, such as color and shape, allows for more nuanced visual differentiation:

```{r}
# Multiple aesthetics - color and shape
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species)) +
  geom_point()
```

### Adding Smoothing

You can add smoothing to visualize trends more clearly. Here’s how to include `geom_smooth()` with different methods:

```{r}
# Smoothing with loess
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species)) +
  geom_point() +
  geom_smooth(method = "loess")
```

```{r}
# Smoothing with gam
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x))
```

### Adding Labels

Enhance your plot by adding titles and axis labels:

```{r}
# Adding labels
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x)) +
  labs(title = "Flipper Length vs. Body Mass by Species",
       x = "Flipper Length (mm)",
       y = "Body Mass (g)")
```

### Adding Size and Alpha

You can further differentiate your data by mapping additional aesthetics such as size and alpha (transparency):

```{r}
# Multiple aesthetics - color, shape, and size
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species, size = species)) +
  geom_point()
```

```{r}
# Alpha aesthetic
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species, size = species, alpha = species)) +
  geom_point()
```

```{r}
# Only alpha
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, alpha = species)) +
  geom_point()
```

### Setting Aesthetic Apart

If you want to set a global aesthetic (e.g., color all points the same), place the aesthetic code outside of `aes()`:

```{r}
# Global aesthetic - color
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(color = "purple")
```

```{r}
# Equivalent
ggplot(penguins) +
  geom_point(aes(x = flipper_length_mm, y = body_mass_g), color = "purple")
```

# Conclusion

We covered common aesthetics in `ggplot2`, including color, shape, size, and alpha, and demonstrated how they can enhance your visualizations. Feel free to revisit these concepts and practice using the `ggplot2` package to make your data visualizations more informative and engaging. 

