---
title: "Module 4: Customizing Plots with Labels and Annotations"
author: "Saswat Seth"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---

 

# Customizing Plots with Labels and Annotations

In this section, we'll learn how to enhance our ggplot2 visualizations by adding titles, subtitles, captions, and annotations. These elements help clarify the purpose of your plot, highlight key points, and make your data more accessible to your audience.

Let's start by loading the necessary libraries.

```{r}
library(ggplot2)
library(palmerpenguins)

data = penguins
```

## Adding a Title

We can use the `labs()` function to add a title to our plot. Titles are important for quickly conveying the purpose of your visualization.

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(title = "Palmer Penguins: Body Mass vs Flipper Length")
```

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(title = "Palmer Penguins: Body Mass vs Flipper Length")

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(
    title = "Palmer Penguins: Body Mass vs Flipper Length",
    subtitle = "Sample of Three Penguin Species"
  )
```

## Adding a Caption

Captions are useful for citing the source of your data or providing any other relevant information. They are displayed at the bottom of the plot.

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(
    title = "Palmer Penguins: Body Mass vs Flipper Length",
    subtitle = "Sample of Three Penguin Species",
    caption = "Data collected by Dr. Kristen Gorman"
  )
```

## Adding Annotations

Annotations allow you to add custom text directly inside your plot, making it easy to call out specific data points. We can use the `annotate()` function for this.

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(
    title = "Palmer Penguins: Body Mass vs Flipper Length",
    subtitle = "Sample of Three Penguin Species",
    caption = "Data collected by Dr. Kristen Gorman"
  ) +
  annotate("text", x = 220, y = 3500, label = "The Gentoos are the largest", 
           color = "purple", fontface = "bold", size = 4.5, angle = 25)
```

## Using Variables to Store Plots

As your code gets longer, it can be useful to store your plot in a variable. This makes it easier to reuse and modify the plot without rewriting the entire code.

```{r}
p <- ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(
    title = "Palmer Penguins: Body Mass vs Flipper Length",
    subtitle = "Sample of Three Penguin Species",
    caption = "Data collected by Dr. Kristen Gorman"
  )

# Printing the plot
p

# Adding an annotation to the stored plot
p + annotate("text", x = 220, y = 3500, label = "The Gentoos are the largest", 
             color = "purple", fontface = "bold", size = 4.5, angle = 25)
```

## Storing the Entire Plot

You can also store the entire customized plot into a variable, making it even more flexible for future use.

```{r}
whole <- ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(
    title = "Palmer Penguins: Body Mass vs Flipper Length",
    subtitle = "Sample of Three Penguin Species",
    caption = "Data collected by Dr. Kristen Gorman"
  ) +
  annotate("text", x = 220, y = 3500, label = "The Gentoos are the largest", 
           color = "purple", fontface = "bold", size = 4.5, angle = 25)

# Printing the stored plot
whole
```

# Conclusion

In this section, we explored how to enhance our ggplot2 visualizations by adding titles, subtitles, captions, and annotations. These features are crucial for making your plots more informative and engaging. By customizing the look and feel of your plots, you can better communicate the story behind your data.

We also learned how to manage long pieces of code by storing plots in variables, making it easier to modify and reuse them. With these tools, you're now better equipped to create clear and effective visualizations that can capture your audience's attention and convey your insights more effectively.

Keep practicing these techniques, and you'll continue to improve your data visualization skills. Happy plotting!

