---
title: "Module 4: Aesthetics and Facets in ggplot2"
author: "Saswat Seth"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---
 

# Introduction

In this lesson, we’ll explore how to use `ggplot2` facet functions to display subsets of data in individual plots. Faceting allows us to break down complex data and focus on relationships between different variables.

# Facet Wrap

The `facet_wrap()` function is used when you want to facet your plot by a single variable. Let's start by creating separate plots for each species of penguin to examine the relationship between flipper length and body mass.

```{r}
# Load the necessary libraries and data
library(ggplot2)
library(palmerpenguins)
```

```{r}
data(penguins)
```

```{r}
# Create a scatter plot faceted by species
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  facet_wrap(~ species)
```

## Adding Color by Species

You can also color the points based on species, making it easier to distinguish between the different groups.

```{r}
# Scatter plot with color aesthetics applied to species
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  facet_wrap(~ species)
```

## Alternative Syntax

You can also specify the color aesthetic inside the `geom_point()` function, which is useful if you plan to add other geoms that should not use the color aesthetic.

```{r}
# Alternative syntax for adding color aesthetics
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species)) +
  facet_wrap(~ species)
```

## Facet Grid

When you want to facet by two variables, use `facet_grid()`. This function will split the plot into facets vertically by the first variable and horizontally by the second.

```{r}
# Scatter plot faceted by both sex and species
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  facet_grid(sex ~ species)
```

## Focusing on One Variable

You can also focus your facet grid on just one of the variables. For example, let's create a grid by species only.

```{r}
# Facet grid by species only
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  facet_grid(~ species)
```

## Applying Facets to Other Data

Faceting is not limited to scatter plots. You can use it with other plot types, such as bar charts. Let's apply faceting to the `diamonds` dataset.

```{r}
# Bar chart faceted by cut
data(diamonds)

ggplot(diamonds, aes(x = cut, fill = cut)) +
  geom_bar() +
  facet_wrap(~ cut)
```

```{r}
# Bar chart faceted by cut with fill based on clarity
ggplot(diamonds, aes(x = cut, fill = clarity)) +
  geom_bar() +
  facet_wrap(~ cut)
```

# Conclusion

Faceting with ggplot2 is a powerful way to organize and display complex data. Whether you're breaking down your data by one variable or multiple variables, facets help you uncover important patterns and relationships that might be hidden in a single plot.

