---
title: An R Markdown document converted from "C:\Users\saswa\Downloads\module-4-getting-started-with-ggplot.ipynb"
output: html_document
---

```{r}
# This R environment comes with many helpful analytics packages installed
# It is defined by the kaggle/rstats Docker image: https://github.com/kaggle/docker-rstats
# For example, here's a helpful package to load

library(tidyverse) # metapackage of all tidyverse packages

# Input data files are available in the read-only "../input/" directory
# For example, running this (by clicking run or pressing Shift+Enter) will list all files under the input directory

list.files(path = "../input")

# You can write up to 20GB to the current directory (/kaggle/working/) that gets preserved as output when you create a version using "Save & Run All" 
# You can also write temporary files to /kaggle/temp/, but they won't be saved outside of the current session
```

# Exploring the Palmer Penguins Data with ggplot2

In this notebook, we will use the Palmer Penguins dataset to create visualizations using `ggplot2`. This dataset contains size measurements for three species of penguins that live in the Palmer Archipelago, Antarctica. We will learn how to plot the relationship between different variables using `ggplot2`.

Before we start, make sure that you have the necessary packages installed and loaded.

## Setup

First, we need to install and load the required packages. We'll use `tidyverse` which includes `ggplot2` for creating plots and `palmerpenguins` for accessing the dataset.

```{r}
install.packages("tidyverse")

library(tidyverse)

install.packages("palmerpenguins")

library(palmerpenguins)
```

## Plotting Body Mass vs. Flipper Length

Let's create a scatter plot to visualize the relationship between body mass and flipper length for the penguins. We'll use `ggplot2` to generate this plot.

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  labs(x = "Flipper Length (mm)", y = "Body Mass (g)", 
       title = "Scatter Plot of Body Mass vs. Flipper Length")
```

## Plotting Bill Length vs. Bill Depth

Now, let's create a scatter plot to visualize the relationship between bill length and bill depth. This will help us understand how these measurements relate to each other.

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill Length (mm)", y = "Bill Depth (mm)", 
       title = "Scatter Plot of Bill Length vs. Bill Depth")
```

## Help and Resources

If you need more information about any `ggplot2` function, you can use the help feature in R. For example, to learn more about `geom_point`, run:

```{r}
??geom_point

??tidyverse
```

## Conclusion

In this notebook, we explored how to use `ggplot2` to visualize data from the Palmer Penguins dataset. We created scatter plots to examine the relationships between body mass and flipper length, as well as bill length and bill depth. 

These visualizations help us understand the correlations between different measurements and provide insights into the characteristics of penguins in the Palmer Archipelago. 

By following the steps demonstrated here—setting up the plot with `ggplot()`, adding layers with `geom_point()`, and mapping variables with `aes()`—you can create various types of plots and explore different aspects of your data.

Feel free to experiment with other variables and geoms, and use the help resources provided to deepen your understanding of `ggplot2` and data visualization techniques. Happy plotting!

