---
title: An R Markdown document converted from "C:\Users\saswa\Downloads\module-4-hoa-visualizing-data-with-ggplot2.ipynb"
output: html_document
---

```{r}
# This R environment comes with many helpful analytics packages installed
# It is defined by the kaggle/rstats Docker image: https://github.com/kaggle/docker-rstats
# For example, here's a helpful package to load

library(tidyverse) # metapackage of all tidyverse packages

# Input data files are available in the read-only "../input/" directory
# For example, running this (by clicking run or pressing Shift+Enter) will list all files under the input directory

list.files(path = "../input")

# You can write up to 20GB to the current directory (/kaggle/working/) that gets preserved as output when you create a version using "Save & Run All" 
# You can also write temporary files to /kaggle/temp/, but they won't be saved outside of the current session
```

# Introduction to Data Visualization with `ggplot2`

In this notebook, we will explore the basics of creating visualizations in R using the `ggplot2` package.The `ggplot2` package allows us to create high-quality, customizable plots based on the grammar of graphics, which is a systematic way of constructing visualizations from basic building blocks.

## Installing and Loading Required Packages

First, we need to install and load the `ggplot2` package along with the `palmerpenguins` dataset, which we will use for our visualizations.

Install ggplot2 and palmerpenguins packages

```{r}
install.packages("ggplot2")
install.packages("palmerpenguins")
```

Load the packages into the R session

```{r}
library(ggplot2)
library(palmerpenguins)
```

## Exploring the Palmer Penguins Dataset

Let's load the `penguins` dataset and take a quick look at its structure.This dataset contains measurements for three species of penguins: Adelie, Chinstrap, and Gentoo,including variables like body mass, flipper length, and bill length.

Load and view the penguins dataset

```{r}
data(penguins)
View(penguins)
```

## Creating a Scatterplot with `ggplot2`

We will create a scatterplot to visualize the relationship between two variables: 
flipper length and body mass of the penguins. 
This plot will help us understand how these two measurements are related across different penguin species.

Create a scatterplot of flipper length vs body mass

```{r}
ggplot(data = penguins) + 
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g))
```

## Alternative Syntax for Mapping Aesthetics

The same plot can be created using a slightly different syntax where the mapping of aesthetics is done inside the `ggplot()` function call. This alternative method allows for a more concise code structure.

Alternative syntax for the same scatterplot

```{r}
ggplot(data = penguins, mapping = aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point()
```

# Conclusion

These examples demonstrate the basic steps of creating a scatterplot using `ggplot2`. As you continue exploring `ggplot2`, you'll learn how to customize and enhance your visualizations further. The `ggplot2` cheat sheet is a great resource to keep handy as you dive deeper into data visualization in R.

