---
title: An R Markdown document converted from "C:\Users\saswa\Downloads\module-2-vectors-and-lists-in-r.ipynb"
output: html_document
---

```{r}
# This R environment comes with many helpful analytics packages installed
# It is defined by the kaggle/rstats Docker image: https://github.com/kaggle/docker-rstats
# For example, here's a helpful package to load

library(tidyverse) # metapackage of all tidyverse packages

# Input data files are available in the read-only "../input/" directory
# For example, running this (by clicking run or pressing Shift+Enter) will list all files under the input directory

list.files(path = "../input")

# You can write up to 20GB to the current directory (/kaggle/working/) that gets preserved as output when you create a version using "Save & Run All" 
# You can also write temporary files to /kaggle/temp/, but they won't be saved outside of the current session
```

# Creating Atomic Vectors in R

In R, vectors are a fundamental data structure that can hold elements of the same type. Here, we will explore how to create different types of atomic vectors.

## Numeric Vector

A numeric vector can hold decimal values.

```{r}
numeric_vector <- c(2.5, 48.5, 101.5)
numeric_vector
```

# Integer Vector

An integer vector contains integer values, denoted by the L suffix.

```{r}
integer_vector <- c(1L, 5L, 15L)
```

# Character Vector

A character vector holds string values.

```{r}
character_vector <- c("Sara", "Lisa", "Anna")
```

# Logical Vector

A logical vector contains boolean values.

```{r}
logical_vector <- c(TRUE, FALSE, TRUE)
```

# Determining the Properties of Vectors

We can check the type and length of vectors and verify if a vector is of a specific type.

# Check Type of Character Vector

```{r}
typeof(c("a", "b"))  # "character"
```

# Check Type of Integer Vector

```{r}
typeof(c(1L, 3L))    # "integer"
```

# Determine the Length of a Vector

```{r}
x <- c(33.5, 57.75, 120.05)
length(x)  # [1] 3
```

# Check if a Vector is a Specific Type

```{r}
x <- c(2L, 5L, 11L)
is.integer(x)  # [1] TRUE

y <- c(TRUE, TRUE, FALSE)
is.character(y)  # [1] FALSE
```

# Naming Vectors

We can assign names to the elements of a vector for better readability.

# Create a Named Vector

```{r}
x <- c(1, 3, 5)
names(x) <- c("a", "b", "c")

# Display named vector
x  # a b c 
   # [1] 1 3 5
```

# Creating Lists

Lists in R can contain elements of different types, including other lists.

# Create a List with Different Types

```{r}
my_list <- list("a", 1L, 1.5, TRUE)
```

# Nested Lists

```{r}
nested_list <- list(list(list(1, 3, 5)))
```

# Determining the Structure of Lists

We can check the structure of lists using the str() function.

# Check Structure of a List

```{r}
str(my_list)
```

# Nested List Structure

```{r}
z <- list(list(list(1, 3, 5)))
str(z)
```

# Naming Lists

We can also name the elements of a list for clarity.

# Create a Named List

```{r}
named_list <- list('Chicago' = 1, 'New York' = 2, 'Los Angeles' = 3)

# Display named list
named_list
```

# Conclusion

In this notebook, we explored the fundamentals of atomic vectors and lists in R. We learned how to create and manipulate various types of vectors, including numeric, integer, character, and logical vectors. By understanding the properties of these vectors, such as their type and length, we can effectively manage and analyze data.

Additionally, we delved into lists, a versatile data structure that can hold elements of different types and even nest other lists. Naming vectors and lists enhances clarity and organization, making it easier to reference and understand the data.

These foundational concepts are essential for effective data analysis and will serve as building blocks for more advanced topics in R programming. Mastery of vectors and lists will enable us to handle complex datasets and perform insightful analyses as we progress in our data analytics journey.

