---
title: "Module 3: Organize Your Data"
author: "Saswat Seth"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---
 

## Introduction

In this lesson, we will learn how to organize and filter data in R using functions like `arrange`, `group_by`, `summarize`, and `filter`. These functions are essential for turning data into actionable insights. We'll continue working with the Palmer Penguins dataset and explore various ways to sort, group, and filter the data.

## Step 1: Load Required Packages

We will start by loading the core `tidyverse` package, which includes all the necessary functions for this lesson.

```{r}
library(tidyverse)
```

## Step 2: Sorting Data with `arrange`

To sort our data by a specific variable, we use the `arrange` function. Let's sort the penguin data by bill length in ascending order:

```{r}
penguins %>% arrange(bill_length_mm)
```

To sort the data in descending order, we simply add a minus sign before the column name:

```{r}
penguins %>% arrange(-bill_length_mm)
```

## Step 3: Saving Sorted Data

To save the sorted data as a new data frame, we name it and apply the `arrange` function. This ensures that the original data remains intact.

```{r}
penguins2 <- penguins %>% arrange(-bill_length_mm)
view(penguins2)
```

## Step 4: Grouping Data with `group_by` and `summarize`

We can group data by a specific column using the `group_by` function. This is often combined with the `summarize` function to calculate summary statistics for each group. Let's group the data by island and calculate the mean bill length for each group:

```{r}
penguins %>% group_by(island) %>% drop_na() %>% summarize(mean_bill_length_mm = mean(bill_length_mm))
```

We can also find the maximum bill length for each island by replacing `mean` with `max`:

```{r}
penguins %>% group_by(island) %>% drop_na() %>% summarize(max_bill_length_mm = max(bill_length_mm))
```

## Step 5: Grouping by Multiple Columns

We can group by multiple columns, such as species and island, and then summarize both the maximum and mean bill lengths:

```{r}
penguins %>% group_by(species, island) %>% drop_na() %>% summarize(max_bl = max(bill_length_mm), mean_bl = mean(bill_length_mm))
```

## Step 6: Filtering Data with `filter`

To narrow down our analysis to a specific group, we can use the `filter` function. For example, let's filter the data to only include Adelie penguins:

```{r}
penguins %>% filter(species == "Adelie")
```

## Conclusion

Being able to organize, group, and filter data is crucial in the data analysis process. These functions in R help make data wrangling more efficient and enable us to extract meaningful insights. In the next lesson, we'll explore functions like `separate`, `unite`, and `mutate` to transform our data even further.

