---
title: "Module 3: Same Data, Different Outcome"
author: "Saswat Seth"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---


## Exploring Anscombe's Quartet

In this section, we will explore Anscombe's quartet, a famous dataset that demonstrates the importance of data visualization. Despite having nearly identical summary statistics, the datasets in Anscombe's quartet can look very different when visualized. We'll use various R packages to summarize and visualize the data, which will help us uncover these differences.

### Install and Load Required Packages

First, we'll install and load the necessary packages. This includes `Tmisc` for accessing the quartet data and `datasauRus` for creating engaging visualizations.

```{r}
install.packages("Tmisc")
install.packages("datasauRus")
```

Load and View Anscombe's Quartet Data
Now we'll load Anscombe's quartet data from the `Tmisc` package and view it. This dataset contains four different sets of x and y values that have nearly identical summary statistics.

```{r}
data("quartet")
View(quartet)
```

## Summarize the Data
Let's calculate the mean, standard deviation, and correlation for each set in the quartet. This summary will show us that, despite having similar summary statistics, the datasets may behave differently when plotted.

```{r}
quartet %>% 
  group_by(set) %>% 
  summarize(mean_x = mean(x), sd_x = sd(x), mean_y = mean(y), sd_y = sd(y), cor_xy = cor(x, y))
```

## Visualize the Data
To understand the differences between the datasets, we'll create scatter plots with linear regression lines. This will help us see how the data points are distributed and how the relationships differ across the sets.

```{r}
ggplot(quartet, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  facet_wrap(~ set)
```

```{r}
ggplot(quartet, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(~ set)
```

## Explore Additional Visualizations with datasauRus
We'll use the `datasauRus` package to generate various creative plots of the Anscombe data. This package includes visualizations such as the famous "datasaurus" and other shapes, which illustrate how data can be represented in different ways.

```{r}
library(datasauRus)

ggplot(datasaurus_dozen, aes(x = x, y = y, colour = dataset)) +
  geom_point() +
  theme_void() +
  theme(legend.position = "none") +
  facet_wrap(~ dataset, ncol = 3)
```

# Conclusion
Through these visualizations, you'll see how Anscombe's quartet can appear different despite having similar statistical summaries. This highlights the importance of visualizing data to uncover insights that summary statistics alone may not reveal.

