---
title: An R Markdown document converted from "C:\Users\saswa\Downloads\module-3-transforming-data.ipynb"
output: html_document
---

```{r}
# This R environment comes with many helpful analytics packages installed
# It is defined by the kaggle/rstats Docker image: https://github.com/kaggle/docker-rstats
# For example, here's a helpful package to load

library(tidyverse) # metapackage of all tidyverse packages

# Input data files are available in the read-only "../input/" directory
# For example, running this (by clicking run or pressing Shift+Enter) will list all files under the input directory

list.files(path = "../input")

# You can write up to 20GB to the current directory (/kaggle/working/) that gets preserved as output when you create a version using "Save & Run All" 
# You can also write temporary files to /kaggle/temp/, but they won't be saved outside of the current session
```

# Transforming Data with R

In this section, we will explore how to transform data in R using the `separate()`, `unite()`, and `mutate()` functions. These functions allow us to break up variables across multiple columns, combine existing columns, or even add new calculated values to a data frame.

## Creating a Data Frame

To start, we'll create a simple data frame that contains employee information, including names and job titles. This data frame will serve as the basis for applying various transformation functions.

```{r}
id <- c(1:10)

name <- c("John Mendes", "Rob Stewart", "Rachel Abrahamson", "Christy Hickman", "Johnson Harper", "Candace Miller", "Carlson Landy", "Pansy Jordan", "Darius Berry", "Claudia Garcia")

job_title <- c("Professional", "Programmer", "Management", "Clerical", "Developer", "Programmer", "Management", "Clerical", "Developer", "Programmer")

employee <- data.frame(id, name, job_title)
```

Printing the entire employee data frame

```{r}
print(employee)
```

Displaying the first few rows of the data frame

```{r}
head(employee)
```

Glimpse of the data structure

```{r}
glimpse(employee)
```

Checking the structure of the data frame

```{r}
str(employee)
```

Skimming through the data without charts

```{r}
library(skimr)
skim_without_charts(employee)
```

## Transforming Data: `separate()`, `unite()`, and `mutate()`

## Separate Columns

Right now, the first and last names are combined into one column. We can use the `separate()` function to split these into separate columns.
Separating the 'name' column into 'first_name' and 'last_name'.

```{r}
separate(employee, name, into = c('first_name', 'last_name'), sep = " ")
```

## Unite Columns

Next, we'll use the `unite()` function to merge the `first_name` and `last_name` columns back into a single `name` column.

```{r}
first_name <- c("John", "Rob", "Rachel", "Christy", "Johnson", "Candace", "Carlson", "Pansy", "Darius", "Claudia")

last_name <- c("Mendes", "Stewart", "Abrahamson", "Hickman", "Harper", "Miller", "Landy", "Jordan", "Berry", "Garcia")

job_title <- c("Professional", "Programmer", "Management", "Clerical", "Developer", "Programmer", "Management", "Clerical", "Developer", "Programmer")

employee1 <- data.frame(id, first_name, last_name, job_title)
```

Merging the 'first_name' and 'last_name' columns into 'name'

```{r}
unite(employee1, 'name', first_name, last_name, sep = " ")
```

## Mutate Data

In addition to separating and merging columns, we can also create new variables in our data frame using the `mutate()` function. For instance, we can add a new column with calculated values.

Using the 'penguins' dataset to add a new column with body mass in kilograms

```{r}
library(palmerpenguins)

penguins %>%
  mutate(body_mass_kg = body_mass_g / 1000) %>%
  select(body_mass_kg)
```

We can also create multiple new variables at once using `mutate()` by adding a comma between the calculations.
Adding multiple new variables: body mass in kg and flipper length in cm

```{r}
penguins %>%
  mutate(body_mass_kg = body_mass_g / 1000, flipper_length_cm = flipper_length_mm / 100) %>%
  select(body_mass_kg, flipper_length_cm)
```

# Conclusion

In this notebook, we've learned how to transform existing data in our tables and how to create new variables using `separate()`, `unite()`, and `mutate()`. These are fundamental tools for data manipulation in R, and as you practice, you might discover new ways to use them effectively in your analyses.

