---
title: "Module 3: Create Your Own Data Frame"
author: "Saswat Seth"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---

 


# Lesson 2: Create your own data frame

Background for this activity
This activity is focused on creating and using data frames in R. A data frame is a collection of columns containing data, similar to a spreadsheet or SQL table. Data frames are one of the basic tools you will use to work with data in R, and you can create data frames from different data sources.

There are three common sources for data:

A package with data that can be accessed by loading that package
An external file like a spreadsheet or CSV that can be imported into R
Data that has been generated from scratch using R code
Wherever data comes from, you will almost always want to store it in a data frame object to work with it. Now, you can start creating and exploring data frames with the code chunks below.

## Step 1: Load packages
Start by installing the required package, tidyverse. If you have already installed and loaded tidyverse in this session, you can skip this step.

```{r}
# Install the tidyverse package
install.packages("tidyverse")
```

Once a package is installed, you can load it by running the library() function with the package name inside the parentheses:

```{r}
# Load the tidyverse package
library(tidyverse)
```

## Step 2: Create a data frame
Sometimes you will need to generate a data frame directly in R. One common way is to create individual vectors of data and then combine them into a data frame using the data.frame() function.

First, create a vector of names by inserting four names into the code block:

```{r}
# Create a vector of names
names <- c("Brad", "Bradlin", "Ted", "Euler")
```

Next, create a vector of ages by adding four ages separated by commas. Make sure the ages are numeric values:

```{r}
# Create a vector of ages
age <- c(25, 30, 35, 40)
```

With these two vectors, you can create a new data frame called people:


```{r}
# Create the data frame
people <- data.frame(names, age)
```

## Step 3: Inspect the data frame
Now that you have this data frame, you can use some different functions to inspect it.

Use the head() function to preview the data. This function returns the first few rows of data:

```{r}
# Preview the data frame
head(people)
```

Other useful functions to summarize or preview your data include str() and glimpse(), which provide summaries of each column:

```{r}
# Summarize the data frame structure
str(people)
```

```{r}
# Glimpse the data frame
glimpse(people)
```

To list the column names in your data frame, use the colnames() function:


```{r}
# List the column names
colnames(people)
```

You can also use mutate() to create a new variable. For example, to calculate each person's age in twenty years:

```{r}
# Add a new variable for age in 20 years
mutate(people, age_in_20 = age + 20)
```

## Step 4: Try it yourself
To get more familiar with creating and using data frames, use the code chunks below to create your own custom data frame.

First, create a vector of any five different fruits:

```{r}
# Create a vector of fruits
fruits<-c("Mango","Pineapple","Banana","Apple","Lichi")
```

Now, create a vector with a number representing your rank for each fruit:


```{r}
# Create a vector of rankings
rank<-c(2,1,3,5,4)
```

Finally, combine the two vectors into a data frame called fruit_ranks:


```{r}
# Combine vectors into a data frame
fruit_ranks<-data.frame(fruits,rank)
fruit_ranks;
```

# Activity Wrap Up
In this activity, you learned how to create data frames, view them with summary functions like head() and glimpse(), and make changes with the mutate() function. You can continue practicing these skills by modifying the code chunks in this notebook or use this code as a starting point in your own projects. As you explore data frames, consider how they are similar and different from the tables you have worked with in other data analysis tools like spreadsheets and SQL. Understanding how to create and work with data frames is an important first step to analyzing data in R.


