---
title: "Module 3: More on R Operators"
author: "Saswat Seth"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---

 

# More on R operators
You might remember that an operator is a symbol that identifies the type of operation or calculation to be performed in a formula. In an earlier video, you learned how to use the assignment and arithmetic operators to assign variables and perform calculations. In this reading, you will review a detailed summary of the main types of operators in R, and learn how to use specific operators in R code. 

## Operators 
In R, there are four main types of operators:
1. Arithmetic
2. Relational 
3. Logical
4. Assignment 
Review the specific operators in each category and check out some examples of how to use them in R code.

### Arithmetic operators
Arithmetic operators let you perform basic math operations like addition, subtraction, multiplication, and division. 

The table below summarizes the different arithmetic operators in R. The examples used in the table are based on the creation of two variables: `x` equals `2` and `y` equals `5`. Note that you use the assignment operator to store these values: 

```{r}
x <- 2
y <- 5 
```

| Operator | Description            | Example Code | Result/ Output |
|----------|------------------------|--------------|----------------|
| +        | Addition               | `x + y`      | `[1] 7`        |
| -        | Subtraction            | `x - y`      | `[1] -3`       |
| *        | Multiplication         | `x * y`      | `[1] 10`       |
| /        | Division               | `x / y`      | `[1] 0.4`      |
| %%       | Modulus (remainder)    | `y %% x`     | `[1] 1`        |
| %/%      | Integer division       | `y %/% x`    | `[1] 2`        |
| ^        | Exponent               | `y ^ x`      | `[1] 25`       |

### Relational operators
Relational operators, also known as comparators, allow you to compare values. Relational operators identify how one R object relates to another—like whether an object is less than, equal to, or greater than another object. The output for relational operators is either `TRUE` or `FALSE` (which is a logical data type, or boolean).

The table below summarizes the six relational operators in R. The examples used in the table are based on the creation of two variables: `x` equals `2` and `y` equals `5`. Note that you use the assignment operator to store these values.

```{r}
x <- 2
y <- 5
```

If you perform calculations with each operator, you get the following results. In this case, the output is boolean: `TRUE` or `FALSE`. Note that the `[1]` that appears before each output is used to represent how output is displayed in RStudio.

| Operator | Description               | Example Code | Result/ Output |
|----------|---------------------------|--------------|----------------|
| <        | Less than                 | `x < y`      | `[1] TRUE`     |
| >        | Greater than              | `x > y`      | `[1] FALSE`    |
| <=       | Less than or equal to     | `x <= 2`     | `[1] TRUE`     |
| >=       | Greater than or equal to  | `y >= 10`    | `[1] FALSE`    |
| ==       | Equal to                  | `y == 5`     | `[1] TRUE`     |
| !=       | Not equal to              | `x != 2`     | `[1] FALSE`    |

### Logical operators
Logical operators allow you to combine logical values. Logical operators return a logical data type or boolean (`TRUE` or `FALSE`). You encountered logical operators in an earlier reading, but here is a quick refresher. 

The table below summarizes the logical operators in R.

| Operator | Description                |
|----------|----------------------------|
| &        | Element-wise logical AND   |
| &&       | Logical AND                |
| \|       | Element-wise logical OR    |
| ||       | Logical OR                 |
| !        | Logical NOT                |


Next, check out some examples of how logical operators work in R code. 

#### Element-wise logical AND (&) and OR (|)
You can illustrate logical AND (`&`) and OR (`|`) by comparing numerical values. Create a variable `x` that is equal to `10`. 

```{r}
x <- 10
```

The AND operator returns `TRUE` only if both individual values are `TRUE`. 

```{r}
x > 2 & x < 12
```

10 is greater than 2 and 10 is less than 12. So, the operation evaluates to `TRUE`. 

The OR operator (`|`) works in a similar way to the AND operator (`&`). The main difference is that just one of the values of the OR operation needs to be `TRUE` for the entire OR operation to evaluate to `TRUE`. Only if both values are `FALSE` will the entire OR operation evaluate to `FALSE`.

Now try an example with the same variable (`x <- 10`): 

```{r}
x > 2 | x < 8
```

10 is greater than 2, but 10 is not less than 8. But since at least one of the values (`10 > 2`) is `TRUE`, the OR operation evaluates to `TRUE`. 

```{r}
!(x < 15)
```

The NOT operation evaluates to `FALSE` because it takes the opposite logical value of the statement `x < 15`, which is `TRUE` (10 is less than 15).

The NOT operation evaluates to `FALSE` because it takes the opposite logical value of the statement `x < 15`, which is `TRUE` (10 is less than 15).

### Assignment operators
Assignment operators let you assign values to variables. 

In many scripting programming languages you can just use the equal sign (`=`) to assign a variable. For R, the best practice is to use the arrow assignment (`<-`). Technically, the single arrow assignment can be used in the left or right direction. But the rightward assignment is not generally used in R code. 

You can also use the double arrow assignment, known as a scoping assignment. But the scoping assignment is for advanced R users, so you won’t learn about it in this reading. 

The table below summarizes the assignment operators and example code in R. Notice that the output for each variable is its assigned value.

| Operator | Description            | Example Code  | Result/ Output |
|----------|------------------------|---------------|----------------|
| <-       | Leftwards assignment   | `x <- 2`      | `[1] 2`        |
| <<-      | Leftwards assignment   | `x <<- 7`     | `[1] 7`        |
| =        | Leftwards assignment   | `x = 9`       | `[1] 9`        |
| ->       | Rightwards assignment  | `11 -> x`     | `[1] 11`       |
| ->>      | Rightwards assignment  | `21 ->> x`    | `[1] 21`       |

The operators you learned about in this reading are a great foundation for using operators in R. 

# Conclusion

In this reading, you've reviewed the four main types of operators in R: arithmetic, relational, logical, and assignment operators. Understanding how these operators work and how to use them in R code is essential for performing a wide range of calculations, comparisons, and logical operations.

By mastering these operators, you'll be well-equipped to manipulate data, make comparisons, and assign values effectively in your R programming journey. As you continue to practice, these operators will become second nature, enabling you to write more efficient and powerful R scripts.

Remember, the key to becoming proficient in R is consistent practice and application. Experiment with different operators, combine them in various ways, and apply them to your own data analysis tasks to solidify your understanding. Happy coding!

