---
title: "Module 1: Fun with R -  Scatter Plots with ggplot2"
author: "Saswat Seth"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---

# Introduction

In this module, we will explore creating scatter plots using the ggplot2 package in R. We'll use the palmerpenguins dataset to visualize the relationship between flipper length and body mass of different penguin species.

# 1. Load Necessary Libraries

```{r}
# Load necessary packages (ggplot2 is part of tidyverse, so it's already loaded)
library(tidyverse)
library(palmerpenguins)
```

# 2. Basic Scatter Plot

Create a basic scatter plot to visualize the relationship between flipper length and body mass.

```{r}
# Create a scatter plot of flipper length vs. body mass
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

# 3. Handle Missing Values

Handle missing values by removing them explicitly before plotting.

```{r}
# Remove rows with missing values
penguins_clean <- na.omit(penguins)

# Create the plot with cleaned data
ggplot(data = penguins_clean, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

# 4. Adding Color and Shape Aesthetics

Enhance the scatter plot by adding color and shape based on species.

```{r}
# Plot with color aesthetics
ggplot(data = penguins_clean, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point(aes(color = species))

# Plot with shape aesthetics
ggplot(data = penguins_clean, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point(aes(shape = species))

# Plot with both color and shape aesthetics
ggplot(data = penguins_clean, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point(aes(shape = species, color = species))
```

# 5. Faceting the Plot

Use faceting to create individual plots for each species.

```{r}
# Facet the plot by species
ggplot(data = penguins_clean, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point(aes(shape = species, color = species)) +
  facet_wrap(~ species)
```

# 6. Adding Titles

Add a title to the faceted plot for better presentation.

```{r}
# Add title to the plot
ggplot(data = penguins_clean, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point(aes(shape = species, color = species)) +
  facet_wrap(~ species) + 
  labs(title = "Palmer Penguins: Body Mass vs. Flipper Length")
```

Conclusion
This module demonstrated how to create and enhance scatter plots using ggplot2, handle missing data, and use faceting to compare different groups within the data.

